---
title: "Replication files for Gutiérrez-Vargas, Vandebroek & Meulders (2023) "
output: github_document
date: '2022-11-15'
---

This repository has all the necesary functions to replicate the results form Gutiérrez-Vargas, Vandebroek & Meulders (2023) forthcoming at the Journal of Choice Modelling. 

The source code is contained in a small R package I developed to store all the necesary functions which is called `MobMixlogit` and can be found in the folde `\src` and can be installed locally by running the R script `\src\install_MobMixlogit.R`

The different parts of the article are organized in different folders:

* `\simulations_replication`: contains the replications from sections 5 and 6.
* `\empirical_application`: contains the replications from section 7.

---------------------------------------


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Small example:   Decision Tree with a Mixed Logit at the end leaves.

Here is a brief illustration on how to implement a Mixed Logit at the end leaves using the MOB algorithm. We start by creating some data that contains a partition.

```{r dgp , message=FALSE}
library(MobMixlogit)
library(partykit)
library(mlogit)
# Simulate some data with a partition (see Figure 2a on the article)
df <- MobMixlogit::dgp_tree(N     = 500, 
                            t     = 10,
                            xi    = 0.5, 
                            delta = 1) 
# Create the choice variable for the wide format
df$choice <- with(df,ave(x = altern * chosen,
                         group = id,
                         FUN = max))

# Delete chosen variable
df$chosen <- NULL ; df$id_choice_situation_of_individual_n <- NULL
# Reshape from long to wide.
df_wide <- stats::reshape(data      = df,
                          idvar     = c("id","id_ind"),
                          timevar   = "altern",
                          v.names   = c("x1", "x2"),
                          direction = "wide",
                          sep = "_")
```

Now we have to create an auxiliar function that will be usefull to run the Mixed Logit model using `mlogit`.

```{r mixl_for_partykit , message=FALSE}
# Create a auxiliar function to make mlogit work together with partykit.
mixl_for_partykit <- function(y,
                              x = NULL,
                              start = NULL,
                              weights = NULL,
                              offset = NULL, ...) {

  #cbind() x and y
  yx <- cbind(y,x)
  #First declare dfidx data
  d <- dfidx::dfidx(data = yx,
             shape = "wide",
             choice = "choice",
             varying = 2:7, ## first if "y", then "x's" and finally "identifiers".
             sep = "_",
             idx = list(c("id", "id_ind")),  #obs= choice situation; id = individual
             idnames = c(NA, "altern"))

  # fit mlogit using d
  clogit_dfidx <- mlogit::mlogit(formula = choice ~ x1 + x2 |0 ,  ##+0 to drop intercepts
                         data    = d,           # Dataset
                         rpar    = c(x1 = 'n',  # Both random parameters are assumed normal
                                     x2 = 'n'), 
                         R       = 500,         # Number of Halton draws
                         panel   = TRUE,        # Activate Panel structure (more than one response per individual)
                         method = 'bhhh')       # Optimization technique.

  return(clogit_dfidx)
}
```

Now we call the `partykit::mob()` function and provide our auxiliary function `mixl_for_partykit` as our model fit. 

```{r fit_mob_mixl , message=FALSE}
mob_mixl <- partykit::mob(formula = choice ~  x1_1 + x1_2 + x1_3 + x2_1 + x2_2 + x2_3 + id + id_ind + 0  | z1 +z2 + z3 + z4 + z5 ,
                        data = df_wide,
                        fit = mixl_for_partykit, #auxliar function just defined above
                        cluster = id_ind,        #cluster standard errors
                        control = partykit::mob_control(
                          ytype = "data.frame",
                          xtype = "data.frame",
                          minsize = 100,    # Minumum number of choice sets to create a partition
                          alpha   = 0.05 )) # Level of confidence to reject null hyphotesis
                                              

mob_mixl
```

Finally, we can also provide a graphical illustration of the estimated tree.

```{r pressure, echo=FALSE}
## Summary function inspired by: 
## https://stackoverflow.com/questions/65495322/partykit-modify-terminal-node-to-include-standard-deviation-and-significance-of/65500344#65500344
fn_summary_for_partykit_plots <- function(info, digits = 2) {
  n <- info$nobs
  individuals <- length(unique(info$object$model$idx$id))
  c(paste("T =", n),
    paste("N =", individuals))
  
}

plot(mob_mixl,
     terminal_panel = node_terminal,
     tp_args = list(FUN = fn_summary_for_partykit_plots,
                      fill = "white"))
```


---------------------------------------

## MOB algorithm as variable selection

In our article we also provided an illustration of one possible way to use the MOB algorithm as a variable selection step for the allocation model of Latent Class models. You can find this code in the script `\simulations_replication\MOB_MNL_on_LCMIXL_data.R`.




