---
title: "Modelling preference heterogeneity using model-based decision trees"
output: github_document
---

This repository has all the necessary functions to replicate the results from the article *"Modelling preference heterogeneity using model-based decision trees"* [(Gutiérrez-Vargas, Vandebroek & Meulders, 2023)](https://www.sciencedirect.com/science/article/pii/S1755534522000501) forthcoming at the [Journal of Choice Modelling](https://www.sciencedirect.com/journal/journal-of-choice-modelling). You can find the In Press pre-print [here](https://www.sciencedirect.com/science/article/pii/S1755534522000501).

The source code is contained in a small R package I developed to store all the necessary functions which is called `MobMixlogit` and can be found in the [`\src`](https://github.com/alvarogutyerrez/mobmixl/tree/main/src) folder and can be installed locally by running the R script [`\src\install_MobMixlogit.R`](https://github.com/alvarogutyerrez/mobmixl/blob/main/src/install_MobMixlogit.R)

The different parts of the article are organized in different folders:

-   [`\simulations_replication`](https://github.com/alvarogutyerrez/mobmixl/tree/main/simulations_replication): contains the replications from sections 5 and 6.
-   [`\empirical_application`](https://github.com/alvarogutyerrez/mobmixl/tree/main/empirical_application): contains the replications from section 7.

------------------------------------------------------------------------

For citations please use:

```
@article{gutierrez_vargas_2023,
title = {Modeling preference heterogeneity using model-based decision trees},
journal = {Journal of Choice Modelling},
volume = {46},
pages = {100393},
year = {2023},
issn = {1755-5345},
doi = {https://doi.org/10.1016/j.jocm.2022.100393},
url = {https://www.sciencedirect.com/science/article/pii/S1755534522000501},
author={Guti{\'e}rrez-Vargas, {\'A}lvaro A and Meulders, Michel and Vandebroek, Martina}
}
```

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Small example: Decision Tree with a Mixed Logit at the end leaves.

Here is a brief illustration on how to implement a Mixed Logit at the end leaves using the MOB algorithm. We start by creating some data that contains a partition.

```{r dgp , message=FALSE}
library(MobMixlogit)
library(partykit)
library(mlogit)
# Simulate some data with a partition (see Figure 2a on the article)
df <- MobMixlogit::dgp_tree(N     = 500, # 500 individuals
                            t     = 10,  # 10 choice sets each
                            xi    = 0.5, # xi = 0.5 implies balanced groups (see Section 5.1)
                            delta = 1    # delta regulates how different are the parameters of each partition  
                            ) 
# Create the choice variable for the wide format
df$choice <- with(df,ave(x = altern * chosen,
                         group = id,
                         FUN = max))

# Delete chosen variable
df$chosen <- NULL ; df$id_choice_situation_of_individual_n <- NULL
# Reshape from long to wide.
df_wide <- stats::reshape(data      = df,
                          idvar     = c("id","id_ind"),
                          timevar   = "altern",
                          v.names   = c("x1", "x2"),
                          direction = "wide",
                          sep       = "_")
```

Now we have to create an auxiliary function that will be useful to run the Mixed Logit model using `mlogit`. This way of constructing the auxiliary function was greatly inspired by an answer given by professor [Achim Zeileis](https://www.zeileis.org/) on [StackOverflow](https://stackoverflow.com/questions/67366612/partykitmob-mlogit-error-no-suitable-fitting-function-specified).

```{r mixl_for_partykit , message=FALSE}
# Create a auxiliar function to make mlogit work together with partykit.
mixl_for_partykit <- function(y,
                              x = NULL,
                              start = NULL,
                              weights = NULL,
                              offset = NULL, ...) {
  # Create an artificial dataset binding the x's and the choice variable (y)
  yx <- cbind(y,x)
  # Declare dfidx data
  d <- dfidx::dfidx(data = yx,
             shape = "wide",
             choice = "choice",
             varying = 2:7, ## first if "y", then "x's" and finally "identifiers".
             sep = "_",
             idx = list(c("id", "id_ind")),  #obs= choice situation; id = individual
             idnames = c(NA, "altern"))

  # Fit a Mixed Logit using mlogit to the "d" data.
  clogit_dfidx <- mlogit::mlogit(formula = choice ~ x1 + x2 |0 ,  #+0 to drop intercepts
                         data    = d,           # dataset
                         rpar    = c(x1 = 'n',  # Both random parameters are assumed normal
                                     x2 = 'n'), 
                         R       = 500,         # Number of Halton draws
                         panel   = TRUE,        # Activate Panel structure (more than one response per individual)
                         method = 'bhhh')       # Optimization technique.

  return(clogit_dfidx)
}
```

Then we call the [`partykit::mob()`](https://rdrr.io/cran/partykit/man/mob.html) function and provide our auxiliary function `mixl_for_partykit` as our model fit to it.

```{r fit_mob_mixl , message=FALSE}
(mob_mixl <- partykit::mob(formula = choice ~      # Choice variable
                            x1_1 + x1_2 + x1_3 +  # Variables for the first attribute
                            x2_1 + x2_2 + x2_3 +  # Variables of the second attribute
                            id + id_ind + 0  |    # Identifiers + drop the intercept.
                            z1 +z2 + z3 + z4 + z5 ,
                        data = df_wide,           # data set to be used (wide format)
                        fit = mixl_for_partykit,  # auxiliary function just defined above
                        cluster = id_ind,         # cluster standard errors by individual
                        control = partykit::mob_control(
                          ytype = "data.frame",   # Declare the choice variable as data.frame()
                          xtype = "data.frame",   # Declare the attribute variables as data.frame()
                          minsize = 100,          # Minimum number of choice sets to create a partition
                          alpha   = 0.05 )))      # Level of confidence to reject null hypothesis
```

Finally, we can also provide a graphical illustration of the estimated tree. The figure shows that a first partition was made using variable Z_1, and then a subsequent partition on Z_2. We observe also that both tests rejected the hypothesis of parameter stability with p-values smaller than 0.001. Some other ways of presenting the end nodes can be found [here](https://stackoverflow.com/questions/65495322/partykit-modify-terminal-node-to-include-standard-deviation-and-significance-of) and [here](https://stackoverflow.com/questions/65734766/generate-table-with-side-by-side-node-models-of-partykitmob-object).

```{r pressure, message=FALSE}
## Summary function inspired by: 
## https://stackoverflow.com/questions/65495322/partykit-modify-terminal-node-to-include-standard-deviation-and-significance-of/65500344#65500344
fn_summary_for_partykit_plots <- function(info, digits = 2) {
  n <- info$nobs
  individuals <- length(unique(info$object$model$idx$id))
  c(paste("T =", n),
    paste("N =", individuals))
  
}

plot(mob_mixl,
     terminal_panel = node_terminal,
     tp_args = list(FUN = fn_summary_for_partykit_plots,
                      fill = "white"))
```

------------------------------------------------------------------------

## MOB algorithm as variable selection

In our article we also provided an illustration of one possible way to use the MOB algorithm as a variable selection step for the allocation model of Latent Class models. You can find this code in the script [`\simulations_replication\MOB_MNL_on_LCMIXL_data.R`](https://github.com/alvarogutyerrez/mobmixl/blob/main/empirical_application/MOB_as_diagnostic_tool/MOB_MNL_diagnostic.R).






